---
title: "Bivariate Linear Regression Analysis on Scope of Practice and Mode Copay for New Patients"
output: github_document
---
# 

```{r}
install.packages("readxl", repos = "https://cran.rstudio.com/")


library(readxl)
df <- read_excel("C:/Users/brean/OneDrive/Documents/PhD/Data/Bivariate anaylsis.xlsx", sheet = 1)

```

## Variables
- **Dependent Variable (DV):** `mode_copay_for_new_patient`
- **Independent Variable (IV):** `Scope of Practice`

```{r}
# Convert Scope of Practice into a binary variable
df$scope_binary <- ifelse(df$`Scope of Practice` == "Full Independent Practice Authority", 1, 0)

table(df$scope_binary)

# Dependent variable

df$DV <- df$mode_copay_for_new_patient

options(repos = c(CRAN = "https://cran.rstudio.com/"))

install.packages("rstanarm", dependencies = TRUE)

install.packages("rstanarm", dependencies = TRUE)
library(rstanarm)

# Bayesian linear regression
model <- stan_glm(DV ~ scope_binary, data = df, family = gaussian())

# Summary
summary(model)

plot(df$scope_binary, df$DV,
     xlab = "Scope of Practice (0 = Restricted, 1 = Full Practice)",
     ylab = "Min Medicare Pricing for New Patients",
     main = "Bivariate Regression: Scope of Practice vs. Pricing",
     pch = 16, col = "black")

abline(model, col = "blue")

coef <- coef(model)
eqn <- paste("y =", round(coef[1], 3), "+", round(coef[2], 3), "* x")

```

The formula DV ~ scope_binary shows that the dependent variable (DV) is being predicted by the independent variable scope_binary.

Sample Size: 4000 total posterior samples were drawn from the model.

Observations: The dataset has 2964 observations, and the model uses 2 predictors (including the intercept).

Intercept: The average value of DV when the predictor (scope_binary) is 0, with a mean estimate of 33.5.

scope_binary: This predictor has an estimated mean of 1.7, with a small standard deviation (0.1), indicating a consistent effect across the samples.
sigma: Represents the standard deviation of the residuals (error term) in the model, which is estimated to be 1.7.

Fit Diagnostics: mean_PPD: This is the mean of the posterior predictive distribution, which helps in assessing how well the model fits the data. It is very close to the intercept estimate (33.7).

Conclusion:
The model appears to have converged well and the results show stable and reasonable estimates. It appears that this model estimates that the predictor scope_binary has a positive effect on the outcome variable DV.



